---
title: "<b> t-tests </b>"
subtitle: "Psychological Research Methods:<br>Data Management & Analysis<br><br> "
author: "Monica Truelove-Hill"
institute: "Department of Clinical Psychology<br>The University of Edinburgh"
date: ""
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: xaringan-themer.css
    nature:
      ratio: '16:9'
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

```{r xaringan-themer, include = FALSE}
library(xaringanthemer)
style_mono_accent(
  base_color = "#18778C",
  header_color = "#000000",
  header_font_google = google_font("Jost"),
  header_font_weight = 500,
  text_font_google = google_font("Jost", "300", "300i", "500", "500i"),
  code_font_google = google_font("Source Code Pro"),
  text_bold_color = '#D95829',
  text_slide_number_color = '#18778C',
  text_font_size = '16pt'
)
```

```{r, echo = F, message = F, warning = F}
library(tidyverse)
knitr::opts_chunk$set(dev = 'svg')

# baseColor <- "#18778C"
# accent1 <- '#D95829'
# accent2 <- '#BF3326'
# accent3 <- '#F29422'

baseColor <- "#007382"
accent1 <- '#008C87'
accent2 <- '#090074'
accent3 <- '#7B007B'

```

## Differences between two means

+ Imagine that you have two groups with these means. 

** Are these groups different?**

.pull-left[
```{r, echo = F}
set.seed(68)
dat <- data.frame(x1=round(rnorm(1000, mean = 20, sd = 2)), y1=round(rnorm(1000, mean = 26, sd = 2)), x2=round(rnorm(1000, mean = 22, sd = 8)), y2=round(rnorm(100, mean = 32, sd = 8)))

basePlot <- ggplot(dat, aes(x1)) + 
  scale_x_continuous(breaks=seq(5, 50, by = 5), limits=c(5, 50)) +
  theme(axis.line.y = element_line(),
        axis.line.x = element_line(),
        panel.grid = element_blank(),
        panel.background = element_blank(),
        axis.ticks = element_blank(),
        axis.text = element_text(size = 14),
        axis.title = element_text(size = 16, face = 'bold')) +
  xlab('x')

basePlot + geom_vline(xintercept = mean(dat$x1), linewidth=1, color = baseColor) +
  geom_vline(xintercept = mean(dat$y1), linewidth=1, color = accent2)
```

]

--

.pull-right[

```{r, echo = F}
basePlot + geom_vline(xintercept = mean(dat$x2), linewidth=1, color = baseColor) +
  geom_vline(xintercept = mean(dat$y2), linewidth=1, color = accent2)
```

]

---

## Differences between two means

** Are these means different?**

.pull-left[
```{r, echo = F, warning=F, message=F}
basePlot + geom_vline(xintercept = mean(dat$x1), linewidth=1, color = baseColor) +
  geom_vline(xintercept = mean(dat$y1), linewidth=1, color = accent2) +
  geom_histogram(fill=baseColor, color = 'gray32', alpha = .8, binwidth = 1) +
  geom_histogram(data = dat, aes(x=y1), fill=accent2, color = 'gray32', alpha = .8, binwidth = 1) +
  theme(axis.title.y = element_blank())
```

]


.pull-right[

```{r, echo = F, warning=F, message = F}
basePlot + geom_vline(xintercept = mean(dat$x2), linewidth=1, color = baseColor) +
  geom_vline(xintercept = mean(dat$y2), linewidth=1, color = accent2) + 
  geom_histogram(data = dat, aes(x=x2), fill=baseColor, color = 'gray32', alpha = .8, binwidth = 1) +
  geom_histogram(data = dat, aes(x=y2), fill=accent2, color = 'gray32', alpha = .8, binwidth = 1) +
  theme(axis.title.y = element_blank())
```

]

---

## Directional vs nondirectional hypotheses

+ Hypotheses may be directional, or make a prediction about the expected direction of the effect
  
  + Treatment X will **reduce** symptoms of anxiety
  
  + Test scores will **improve** as hour of study increase

+ Hypotheses can also be nondirectional. In this case, an effect is predicted, but the direction is not specified.

  + Treatment X will **have an effect** on symptoms of anxiety
  
  + Test scores will **be associated with** hours of study

--

+ Directional hypotheses can be tested with one-tailed tests

+ Nondirectional hypotheses can be tested with two-tailed tests

---
exclude: true

## Stating your Hypothesis

---
exclude: true

## Effect size - Cohen's $d$

$$d = \frac{\bar{x}_1 -\bar{x}_2}{s}$$

.pull-left[
+ $\bar{x}_1$ = mean of group 1
+ $\bar{x}_2$ = mean of group 2
+ $s_{pooled}$ = pooled standard deviation
]

--

.pull-right[

$$s_{pooled} = \sqrt{\frac{(n_1-1)s_1^2\ + (n_2-1)s_2^2}{n_1+n_2-2}}$$

]


---
exclude: true
## Effect size - Cohen's $d$

+ Works best for larger samples...for smaller samples (<50), it overinflates results **MONICA, VERIFY THIS**
---
exclude: true
## Power

---
exclude: true
## Assumptions of $t$-test