---
title: "<b> t-tests </b>"
subtitle: "Psychological Research Methods:<br>Data Management & Analysis<br><br> "
author: "Monica Truelove-Hill"
institute: "Department of Clinical Psychology<br>The University of Edinburgh"
date: ""
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: xaringan-themer.css
    nature:
      ratio: '16:9'
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

```{r xaringan-themer, include = FALSE}
library(xaringanthemer)
style_mono_accent(
  base_color = "#18778C",
  header_color = "#000000",
  header_font_google = google_font("Jost"),
  header_font_weight = 500,
  text_font_google = google_font("Jost", "300", "300i", "500", "500i"),
  code_font_google = google_font("Source Code Pro"),
  text_bold_color = '#D95829',
  text_slide_number_color = '#18778C',
  text_font_size = '16pt'
)
```

```{r, echo = F, message = F, warning = F}
library(tidyverse)
knitr::opts_chunk$set(dev = 'svg')

baseColor <- "#18778C"
accent1 <- '#D95829'
accent2 <- '#BF3326'
accent3 <- '#F29422'

```


## Directional vs nondirectional hypotheses

+ Hypotheses may be directional, or make a prediction about the expected direction of the effect
  
  + Treatment X will **reduce** symptoms of anxiety
  
  + Test scores will **improve** as hour of study increase

+ Hypotheses can also be nondirectional. In this case, an effect is predicted, but the direction is not specified.

  + Treatment X will **have an effect** on symptoms of anxiety
  
  + Test scores will **be associated with** hours of study

--

+ Directional hypotheses can be tested with one-tailed tests

+ Nondirectional hypotheses can be tested with two-tailed tests