---
title: "<b> Null Distributions, Effect Sizes, & Power </b>"
subtitle: "Psychological Research Methods:<br>Data Management & Analysis<br><br> "
author: "Monica Truelove-Hill"
institute: "Department of Clinical Psychology<br>The University of Edinburgh"
date: ""
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: xaringan-themer.css
    nature:
      ratio: '16:9'
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

```{r xaringan-themer, include = FALSE}
library(xaringanthemer)
style_mono_accent(
  base_color = "#18778C",
  header_color = "#000000",
  header_font_google = google_font("Jost"),
  header_font_weight = 500,
  text_font_google = google_font("Jost", "300", "300i", "500", "500i"),
  code_font_google = google_font("Source Code Pro"),
  text_bold_color = '#4CA384',
  text_slide_number_color = '#18778C',
  text_font_size = '16pt'
)
```

```{r, echo = F, message = F, warning = F}
library(tidyverse)
knitr::opts_chunk$set(dev = 'svg')

baseColor <- '#4CA384'
accent1 <- '#9AD079'
accent2 <- '#18778C'
accent3 <- '#19424C'
```

## This Week's Topics

+ Samples vs Populations

+ The Null Distribution

+ p-values and critical values

+ Confidence Intervals

+ Effect Sizes

---
## Refresh from Last Week

.pull-left[

+ Probability distributions allow you to visualise the probability of specific observations

+ Specifically, we can compute the probability of a specific range of values within a distribution
]


.pull-right[

```{r, echo = F, fig.height=3.5}
ggplot(data.frame(x = c(-4, 4)), aes(x = x)) +
  stat_function(fun=dnorm, geom = "line", linewidth = 1.5) + 
  theme(axis.text = element_text(size = 12), 
        axis.title = element_text(size = 14, face = 'bold'),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank()) +
  geom_hline(yintercept = 0, linewidth = 1.5)

```
]

---

## Describing your data: Probability distributions

.pull-left[

+ We can also use probability distributions to compute the probability of value beyond a specific point on the distribution. 

+ Imagine we have a range of observations whose values fall between 8 and 32 that are distributed as pictured.

+ Let's say we randomly select an observation from our dataset.

]

.pull-right[
```{r, echo = F, fig.height=3.5}
testDat <- ggplot(data.frame(x = c(8, 32)), aes(x = x)) +
  stat_function(fun=dnorm, args= list(20, 3), geom = "line", linewidth = 1.5) +
  theme(axis.text = element_text(size = 12), 
        axis.title = element_text(size = 14, face = 'bold'),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank()) +
  geom_hline(yintercept = 0, linewidth = 1.5)

testDat
```
]

---
## Describing your data: Probability distributions

.pull-left[

+ We can also use probability distributions to compute the probability of value beyond a specific point on the distribution.

+ Imagine we have a range of observations whose values fall between 8 and 32 that are distributed as pictured.

+ Let's say we randomly select an observation from our dataset.

+ What is the likelihood our randomly selected observation is at least 25?

]

.pull-right[

```{r, echo = F, fig.height=3.5}
testDat + 
    geom_vline(xintercept = 25, color = accent3, linewidth = 1.5)
```

]


---
count: false

## Describing your data: Probability distributions

.pull-left[

+ We can also use probability distributions to compute the probability of value beyond a specific point on the distribution.

+ Imagine we have a range of observations whose values fall between 8 and 32 that are distributed as pictured.

+ Let's say we randomly select an observation from our dataset.

+ What is the likelihood our randomly selected observation is at least 25?

+ ~95% of our observations are less than 25. This means we only have about a 5% chance of randomly selecting an observation of at least 25. 

]

.pull-right[

```{r, echo = F, fig.height=3.5}
testDat + 
  geom_area(stat = 'function', fun = dnorm, args= list(20, 3), fill = baseColor, xlim=c(8, 25), alpha = .8) +
  geom_area(stat = 'function', fun = dnorm, args= list(20, 3), fill = accent2, xlim=c(25, 30), alpha = .8) +
  annotate('text', label = '95%', x = 20, y = .05, color = 'white', size = 7) + 
  annotate('text', label = '5%', x = 26, y = .01, color = 'white', size = 7) +
  geom_vline(xintercept = 25, color = accent3, linewidth = 1)
```
]

---
## Probability Distributions

+ Why does this matter?

--

+ Many of the statistical tests we will learn about in this course all involve testing the probability of a specific value of a test statistic, given that $H_0$ is true. 

+ In other words, if $H_0$ is true, how likely is it that we would get results at least as extreme as ours?

+ If the value is unusual enough (i.e., if it is highly unlikely we would see the value if the null is true), we reject $H_0$


---
## Null Distribution

+ All parametric quantitative tests we learn about in this course evaluate significance by evaluating results against a predefined **null distribution**

+ The null distribution is the probability distribution of a statistic **given that $H_0$ is true**. 

+ The shape of the null distribution depends on the test statistic (and the degrees of freedom, which we'll get to at a later point in the course)


---
## Comparing Data against null distribution: NHST

.pull-left[
```{r, echo = F, fig.height=5}
testDat
```

]


---
count: false

## Comparing Data against null distribution: NHST

.pull-left[

```{r, echo = F, fig.height=5}
testDat + geom_vline(xintercept = 24, color = accent1, linewidth = 1.5)
```

]

.pull-right[
```{r, echo = F, fig.height=5}
testDat + geom_vline(xintercept = 18, color = accent3, linewidth = 1.5)
```

]

> **Test Your Understanding:** Which score is less common?

---
count: false

## Comparing Data against null distribution: NHST

.pull-left[
```{r, echo = F, fig.height=5}
testDat +
  geom_area(stat = 'function', fun = dnorm, args= list(20, 3), fill = baseColor, xlim=c(8, 24), alpha = .8) +
  geom_area(stat = 'function', fun = dnorm, args= list(20, 3), fill = accent2, xlim=c(24, 30), alpha = .8) +
  annotate('text', label = '91%', x = 18, y = .02, color = 'white', size = 7) + 
  annotate('text', label = '9%', x = 25, y = .02, color = 'white', size = 7)+ 
  geom_vline(xintercept = 24, color = accent1, linewidth = 1.5)
```
]

.pull-right[
```{r, echo = F, fig.height=5}
testDat +
  geom_area(stat = 'function', fun = dnorm, args= list(20, 3), fill = accent2, xlim=c(8, 18), alpha = .8) +
  geom_area(stat = 'function', fun = dnorm, args= list(20, 3), fill = baseColor, xlim=c(18, 30), alpha = .8) +
  annotate('text', label = '25%', x = 16, y = .02, color = 'white', size = 7) + 
  annotate('text', label = '75%', x = 22, y = .02, color = 'white', size = 7) + 
  geom_vline(xintercept = 18, color = accent3, linewidth = 1.5)
```
]

> **Test Your Understanding:** Which score is less common?


---

## Null Distributions

.pull-left[
```{r, echo = F, fig.height=2.5, fig.width=4.5}
ggplot(data.frame(x=c(0, 10)), aes(x=x)) +
  stat_function(fun=dchisq, geom='line', args=c(df = 1), color = baseColor, linewidth=1.5) +
  stat_function(fun=dchisq, geom='line', args=c(df = 2), color = accent1, linewidth=1.5) +
  stat_function(fun=dchisq, geom='line', args=c(df = 3), color = accent2, linewidth=1.5) +
  stat_function(fun=dchisq, geom='line', args=c(df = 4), color = accent3, linewidth=1.5) +
  scale_y_continuous(limits=c(0, .52), breaks = seq(0, .5, .1)) +
  annotate('text', label = 'df = 1', color = baseColor, x = 8, y = .45, size = 5) +
  annotate('text', label = 'df = 2', color = accent1, x = 8, y = .38, size = 5) +
  annotate('text', label = 'df = 3', color = accent2, x = 8, y = .31, size = 5) +
  annotate('text', label = 'df = 4', color = accent3, x = 8, y = .24, size = 5) +
  labs(x = expression(chi^2)) +
  theme(axis.text.y = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.x = element_text(size = 16, face = 'bold'),
        axis.text.x = element_text(size = 14))

```
]


.pull-right[
```{r, echo = F, fig.height=2.5, fig.width=4.5}
ggplot(data.frame(x=c(-6, 6)), aes(x=x)) +
  stat_function(fun=dt, geom='line', args=c(df = 5), color = baseColor, linewidth=1.5) +
  stat_function(fun=dt, geom='line', args=c(df = 10), color = accent1, linewidth=1.5) +
  stat_function(fun=dt, geom='line', args=c(df = 20), color = accent2, linewidth=1.5) +
  stat_function(fun=dt, geom='line', args=c(df = 50), color = accent3, linewidth=1.5) +
  annotate('text', label = 'df = 5', color = baseColor, x = 3, y = .4, size = 5) +
  annotate('text', label = 'df = 10', color = accent1, x = 3, y = .35, size = 5) +
  annotate('text', label = 'df = 20', color = accent2, x = 3, y = .3, size = 5) +
  annotate('text', label = 'df = 50', color = accent3, x = 3, y = .25, size = 5) +
  labs(x = 't-statistic') +
  theme(axis.text.y = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.x = element_text(size = 16, face = 'bold'),
        axis.text.x = element_text(size = 14))

```
]

.center[
```{r, echo = F, fig.height=2.5, fig.width=4.5}
ggplot(data.frame(x=c(0, 5)), aes(x=x)) +
  stat_function(fun=df, geom='line', args=c(df1 = 2, df2=10), color = baseColor, linewidth=1.5) +
  stat_function(fun=df, geom='line', args=c(df1 = 5, df2=10), color = accent1, linewidth=1.5) + 
  stat_function(fun=df, geom='line', args=c(df1 = 5, df2=100), color = accent2, linewidth=1.5) +
  stat_function(fun=df, geom='line', args=c(df1 = 10, df2=100), color = accent3, linewidth=1.5) +
  annotate('text', label = 'df1 = 2, df2 = 10', color = baseColor, x = 4, y = .75, size = 5) +
  annotate('text', label = 'df1 = 5, df2 = 10', color = accent1, x = 4, y = .6, size = 5) +
  annotate('text', label = 'df1 = 5, df2 = 100', color = accent2, x = 4, y = .45, size = 5) +
  annotate('text', label = 'df1 = 10, df2 = 100', color = accent3, x = 4, y = .3, size = 5) +
  labs(x = 'F-statistic') +
  theme(axis.text.y = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.x = element_text(size = 16, face = 'bold'),
        axis.text.x = element_text(size = 14))

```
]

---
class: center, inverse, middle

## Questions?


---
