---
title: "<b> Intro to the Linear Model </b>"
subtitle: "Psychological Research Methods:<br>Data Management & Analysis<br><br> "
author: "Monica Truelove-Hill"
institute: "Department of Clinical Psychology<br>The University of Edinburgh"
date: ""
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: xaringan-themer.css
    nature:
      ratio: '16:9'
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---


```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

```{r xaringan-themer, include = FALSE}
library(xaringanthemer)
style_mono_accent(
  base_color = "#18778C",
  header_color = "#000000",
  header_font_google = google_font("Jost"),
  header_font_weight = 500,
  text_font_google = google_font("Jost", "300", "300i", "500", "500i"),
  code_font_google = google_font("Source Code Pro"),
  text_bold_color = '#4CA384',
  text_slide_number_color = '#18778C',
  text_font_size = '16pt'
)
```

```{r, echo = F, message = F, warning = F}
library(tidyverse)
knitr::opts_chunk$set(dev = 'svg')

# baseColor <- "#18778C"
# accent1 <- '#D95829'
# accent2 <- '#BF3326'
# accent3 <- '#F29422'

baseColor <- '#4CA384'
accent1 <- '#9AD079'
accent2 <- '#18778C'
accent3 <- '#19424C'
```

## This Week's Key Topics

+ Understand what is meant by model

+ Understand key features of linear model

+ Understand what is meant by residuals

+ Understand key principles of least squares

---
## Models & Functions

+ **Model:** a formal representation of the world, or an idea about the way the world is

--

+ Suppose I have a model for the MODEL EXAMPLE HERE

+ I'm using this model to formally represent the relationship between the variables V1 HERE and V2 HERE

---

## Models in Research

+ When you formulate a research aim or make a hypothesis, you're pre-supposing a model. 

+ When you perform statistical testing, you are evaluating that model.

---
## Features of a Line

.pull-left[

Two key features of a line:

+ **Intercept:** The point at which the line crosses the $y$-axis ($x$ = 0)

+ **Slope:** The line's rate of change (steepness)

]

.pull-right[

MONICA - BASIC LINE PLOT HERE
```{r, echo = F}

```

]

---
## Features of a Line

.pull-left[

.center[**Different Intercepts**]
```{r, echo = F}

```


]

.pull-right[

.center[**Different Slopes**]
```{r, echo = F}

```

]