---
title: "<b> Nonparametric Tests </b>"
subtitle: "Psychological Research Methods:<br>Data Management & Analysis<br><br> "
author: "Monica Truelove-Hill"
institute: "Department of Clinical Psychology<br>The University of Edinburgh"
date: ""
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: xaringan-themer.css
    nature:
      ratio: '16:9'
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

## Spearman's Correlation

+ Spearman's $\rho$ (or rank-order correlation) uses data on the rank-ordering of $x$, $y$ responses for each individual.
  
+ Spearman's $\rho$ is a nonparametric version of Pearson's $r$, so it doesn't require the same constraints on your data

+ When would we choose to use the Spearman correlation?
	+ If our data are naturally ranked data (e.g. a survey where the task is to rank foods and drinks in terms of preference)
	+ Our data are ordinal (e.g., Likert Scales)
	+ If the data are non-normal or skewed
	+ If the data shows evidence of non-linearity

---
## Spearman's Correlation

+ Spearman's is not testing for linear relations, it is testing for increasing monotonic relationship.

--

  + What?

--

.pull-left.center[
**Perfectly Linear Relationship**
```{r, echo = F, fig.height=3, fig.width=5}
mono <- data.frame(
  A = c(1,2,3,4,5,6,7,8),
  B = c(1,2,3,4,5,6,7,8),
  C = c(1,4,5,6,8,9,10,13)
)

lp <- mono %>%
  ggplot(., aes(x=A, y=B)) +
  geom_point() + 
  geom_line()

mp <- mono %>%
  ggplot(., aes(x=A, y=C)) +
  geom_point() + 
  geom_line()

lp
```

]

.pull-right.center[
**Perfectly Increasing Monotonic Relationship**
```{r, echo = F, fig.height=3, fig.width=5}
mp
```

]

---
## Monotonic Relationship

+ **Perfect Monotonic Relationship:** The rank position of all observations on Variable A is the same as the rank position of all observations on Variable C.

.pull-left[

```{r, echo = F}
mono %>%
  mutate(
    ID = paste("ID", 1:8, sep = ""),
    Rank_A = rank(A),
    Rank_C = rank(C)
  ) %>%
  dplyr::select(., ID, A, C, Rank_A, Rank_C) %>%
  kable() %>%
  kable_styling(full_width = F)

```

]

.pull-right[
```{r, echo = F, fig.height=3.5, fig.width=5}
mp
```

]

---

## Calculating Spearman's $\rho$

$$\rho=1-\frac{6\sum{d^2_i}}{n(n^2-1)}$$
+ $d_i=$ rank( $x_i$ ) $-$ rank( $y_i$ )

+ Steps
  1. Rank each variable from largest to smallest 
  2. Calculate the difference in rank for each person on the two variables
  3. Square the difference
  4. Sum the squared values
