---
title: "Week 9: Regression II"
output:
  learnr::tutorial:
    progressive: true
    allow_skip: true
    css: https://mtruelovehill.github.io/PRM/Labs/css/style.css
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
# library(learnr)
# library(tidyverse)
# library(kableExtra)
# library(plotly)
# library(shiny)

baseColor <- '#4CA384'
accent1 <- '#9AD079'
accent2 <- '#C4C6C7'
accent3 <- '#19424C'
```


## Intro to Today's Lab

During today's lab, you'll be working through the formative assessment. This assessment does not contribute to your course mark. It serves only as a check-in so that you are more aware of what areas you're confident in, and which areas you may need to revise further. The tutors will be available to clarify any feedback or provide further feedback as necessary. Good luck!


## Making a hypothesis

```{r, echo = F}
dat <- read.csv('https://mtruelovehill.github.io/PRM/Labs/Week8LabData.csv')

datInfo <- data.frame(VariableName = colnames(dat),
                      Description=c('Age in years',
                                    'Levels: Female; Male; Nonbinary; Other; Prefer not to Disclose',
                                    'Amount of physical activity per week in hours',
                                    'Amount of sleep per week in hours',
                                    'Scores on a social support scale; values may range between 0-45',
                                    'Scores on a well-being scale; values may range between 0-40'))

datInfo %>%
  kbl(col.names=c('Variable Name', 'Description')) %>%
  kable_styling(full_width = F) %>%
  row_spec(0, bold = T, color=baseColor, font_size = 18, align='l') %>%
  column_spec(1, bold = T, width = '4.5cm')
```


## Selecting an Appropriate Test


```{r cyu1, echo = FALSE}
quiz(caption = '',
     question("A researcher wants to know whether her sample’s scores on a personality test are significantly different than 50. Which of the following statistical tests should she use?",
              answer('One-Sample t-test', correct = T),
              answer('independent-samples t-test'),
              answer('paired-samples t-test'),
              answer('One-Way ANOVA'),
              answer('Factorial ANOVA'),
              answer('Simple Linear Regression'),
              answer('Multiple Linear Regression'),
              message = "The researcher is comparing her sample's mean to a single, pre-established value. A one-sample t-test would be the best method for testing this."),
     question('At what level of measurement does the dependent variable need to be for a t-test?',
              answer('Continuous', correct = T),
              answer('Nominal'),
              answer('Ordered Categories'),
              answer('Binary'),
              message = 'A t-test requires the dependent variable to be continuous.',
              random_answer_order = T))
```


## Checking your Data

## Checking Assumptions

## Running Statistical Tests


## Interpreting Results

Monica - lots of questions with images of SPSS output - which of the following interpretations is correct? that sort of thing.

## Power Analysis

## Interpreting Effect Sizes?


## Reporting?

