---
title: "Week 6 Lab: ANOVA"
output:
  learnr::tutorial:
    progressive: true
    allow_skip: true
    css: css/style.css
runtime: shiny_prerendered
---


```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(kableExtra)
library(knitr)
library(shiny)
library(shinyjs)
library(bslib)

baseColor <- '#4CA384'
accent1 <- '#9AD079'
accent2 <- '#C4C6C7'
accent3 <- '#19424C'
#tutorial_options(exercise.timelimit = 60)
```

```{r, context = 'server'}
output$distPlot <- renderPlot({
  set.seed(806)
      dat <- data.frame(X=rnorm(300, mean = input$mean, sd = input$sd))
      ggplot(data.frame(X = c(-30, 30)), aes(x = X)) +
        geom_histogram(data=dat, aes(x=X, y = after_stat(density)), fill = '#4CA384', color = '#19424C',
                       binwidth = 2) +
        stat_function(fun=dnorm, geom = "line", args = list(mean = input$mean, sd = input$sd),
                      linewidth = 1, color = '#19424C') +
        scale_x_continuous(breaks = seq(-30, 30, 10), limits=c(-30, 30)) +
        scale_y_continuous(limits=c(0, .4)) +
        theme(axis.title.x = element_text(size = 16, face = 'bold'),
              axis.text.x = element_text(size = 16),
              axis.line.y = element_blank(),
              axis.text.y = element_blank(),
              axis.ticks = element_blank(),
              axis.title.y = element_blank())
    })
```

## Intro to Today's Lab


Intro Goes Here   

```{r, echo = F}
# dat <- read.csv('Week1LabData.csv')
# datInfo <- data.frame(VariableName=colnames(dat), 
#                       Description=c('Participant ID', 'Age in years', 'Time of class in which assessment was given', 'Levels: 1=Male,
#                                     2=Females, 3=Nonbinary', 'Assessment Version A Scores', 'Assessment Version B Scores', 'Assessment 
#                                     Version C Scores', 'Assessment Version D Scores'))
# 
# datInfo %>%
#   kbl(col.names=c('Variable Name', 'Description')) %>%
#   kable_styling(full_width = F) %>%
#   row_spec(0, bold = T, color=baseColor, font_size = 18, align='l') %>%
#   column_spec(1, bold = T, width = '4.5cm')
```

## Intro to Today's Lab

During today's lab, you'll apply the concepts discussed during this week's lecture. Each lab consists of a range of tasks, with corresponding questions you can answer. Please note that the questions are not required and not marked, although they do provide a helpful source of formative feedback that will help you gauge your understanding. 

In this week's lab, you'll be working with data simulated based on [this paper](https://d1wqtxts1xzle7.cloudfront.net/92405721/vrtherapy-libre.pdf?1665701596=&response-content-disposition=inline%3B+filename%3DVirtual_Reality_Exposure_Therapy_for_the.pdf&Expires=1723672793&Signature=ceCWXs2EIQsAqYQSz5XM5-rSYxHeEbUHIDYv8cXqDjMrFK6~6PO-5bD05ku0EWCSZaCFaKHER1atUcfjFGUASRZ6L5aij5dBBtF~HV7f3QH70Diq~2P4MXCA5KMJYq9tWPWv4My7b~T5vLnSWGHDI9XYFfEqg7dMPBibKUtaiVTmVekzvXVQwt~E4m0cuVcA1j6KLe3nC6kfGRbO20EMfVZeCLW6f-3M5K3gr~xPnnQBjX7V3m97pJ6FoY8S9SbJzJQiyIGYe1DIh0BIYx-cV7O-a9D~Je8fQrg1wNmBhWdTLJu4h6m3SIop9B-HSTIcW0LmHlxp2SltefXxHSRukA__&Key-Pair-Id=APKAJLOHF5GGSLRBV4ZA). In this study, researchers tested the effectiveness of a virtual reality-based exposure therapy to treat PTSD in victims of the September 11th attacks. Participants were given the Clinician-Administered PTSD Scale (CAPS) and assigned to either a waitlist (n = 8) or treatment group (n = 10). After the treatment group completed therapy, all participants were once again administered the CAPS.


You can download the data for this week's lab [here.](https://mtruelovehill.github.io/PRM/Labs/Week6LabData.csv)

This dataset contains the following variables:

```{r, echo = F}
dat <- read.csv('https://mtruelovehill.github.io/PRM/Labs/Week6LabData.csv')
datInfo <- data.frame(VariableName=colnames(dat), 
                      Description=c('Study Group', 'Age in years', 'Sex', 'pre-intervention CAPS score', 'post-intervention CAPS score'))

datInfo %>%
  kbl(col.names=c('Variable Name', 'Description')) %>%
  kable_styling(full_width = F) %>%
  row_spec(0, bold = T, color=baseColor, font_size = 18, align='l') %>%
  column_spec(1, bold = T, width = '4.5cm')
```



### Learning Objectives
At the end of this lab, you will be able to:

1.  


## State your Hypotheses

Task 1 Info Here

* Identify IV(s) and DV(s)
* State Hypotheses (Mixed Methods ANOVA)

```{r q1, echo = FALSE}
# quiz(caption = 'Test Your Understanding',
#      question('Which of the following hypotheses can be tested using a mixed methods ANOVA? Please select all that apply.',
#               answer('Hypothesis1', correct = T),
#               answer('Hypothesis2'),
#               answer('Hypothesis3'),
#               answer('Hypothesis4'),
#               answer('Hypothesis5'),
#               random_answer_order = T,
#               message = 'Feedback Here'),
#      question('You have developed an intervention and want to test its effectiveness. You recruit a sample of participants and measure their symptom severity both before and after the intervention. You measure patient symptom severity on a continuous scale. You would like to compare the two symptom severity scores. Which of the following tests is most appropriate?',
#               answer('One-sample t-test'),
#               answer('Independent samples t-test'),
#               answer('Paired-samples t-test', correct = T),
#               message = "Because you have a pre- and post-intervention measurement from each participant, the two scores are considered to be paired. An individual's score on the post-test will not be independent of their scores on the pre-test. A paired-samples t-test will account for this lack of independence between the two groups of scores."))

```

<details><summary><span style = "font-weight: bold; font-size: 16pt"> + Click here for a hint </span></summary>

Hint Goes Here

</details>

</br>

<details><summary><span style = "font-weight: bold; font-size: 16pt"> + Click here for the solution </span></summary>

Solution Goes Here

</details>

</br>

## Run the Appropriate Power Analyses 

Task Goes Here
* Maybe an a priori & a sensitivity

```{r q2, echo = FALSE}
# quiz(caption = 'Test Your Understanding',
#      question("Which type of data is C?",
#               answer("Categorical"),
#               answer("Continuous", correct = T),
#               random_answer_order = T),
#      question("Which type of data is Gender?",
#               answer("Continuous"),
#               answer('Categorical', correct = T)),
#      question("Which type of data is ClassTime?",
#               answer("Continuous"),
#               answer('Categorical', correct = T),
#               message='Categorical data are nominal or ordinal data with no numeric relationship (although they may be represented
#               numerically), while continuous data have a traditional numeric relationship; continuous data can technically take on an  
#               infinite range of values if measurement was precise enough'),
#      question('Which of the following "Measure" values used by SPSS reflects categorical data?',
#               answer('Nominal', correct = T),
#               answer('Ordinal', correct = T),
#               answer('Scale'),
#               message = 'Scale can be used for both types of continuous data (ratio and interval scales of measurement), but categorical
#               variables are split into nominal and ordinal measures. Remember that nominal data reflects purely independent categories,
#               but ordinal data has some sort of implied order (e.g. Likert Scale data, level of education).'))
```

<details><summary><span style = "font-weight: bold; font-size: 16pt"> + Click here for a hint </span></summary>

Hint Here

</details>

</br>

<details><summary><span style = "font-weight: bold; font-size: 16pt"> + Click here for the solution </span></summary>

Solution here.

</details>

</br>


## Calculate Descriptive Data 

Task Description Here

<details><summary><span style = "font-weight: bold; font-size: 16pt"> + Click here for a hint </span></summary>

Hint Here

</details>

</br>

<details><summary><span style = "font-weight: bold; font-size: 16pt"> + Click here for the solution </span></summary>

Solution Here.

</details>

</br>




## Visualise Your Data 

Task Here.

```{r q4, echo = FALSE}
# quiz(caption = 'Test Your Understanding',
#      question("At least one of the groups is limited in size compared to the others",
#               answer("True", correct = T),
#               answer("False"),
#               message = 'The nonbinary group is much smaller than the other two measured groups.'),
#      question_numeric("How many total datapoints do you have for the gender variable?",
#               answer(48, correct=T)),
#      question_numeric("What percentage of participants are female?",
#               answer(41.7, correct=T)),
#      question_numeric('How many nonbinary participants are in your sample?',
#               answer(2, correct=T)))
```

<details><summary><span style = "font-weight: bold; font-size: 16pt"> + Click here for a hint </span></summary>

Hint Here

</details>
</br>

<details><summary><span style = "font-weight: bold; font-size: 16pt"> + Click here for the solution </span></summary>

Solution Here

</details>
</br>

## Check Assumptions

Task Here

<details><summary><span style = "font-weight: bold; font-size: 16pt"> + Click here for a hint </span></summary>

Hint Here.

</details>
</br>

<details><summary><span style = "font-weight: bold; font-size: 16pt"> + Click here for the solution </span></summary>

Solution Here.

</details>
</br>

## Run ANOVA

Task Description.
* Maybe also discuss degrees of freedom here

```{r q6, echo = FALSE}
# quiz(caption = 'Test Your Understanding',
#      question("Which of the following central tendency measures could be used to describe age? Please select all that apply.",
#               answer("Mean", correct = T),
#               answer("Median", correct = T),
#               answer('Mode', correct = T),
#               message = 'All of the above could technically be used to describe the Age variable. However, it is most common to use the
#               mean, as it includes information from each observation.'),
#      question("Which of the following central tendency measures could be used to describe gender? Please select all that apply.",
#               answer("Mean"),
#               answer("Median"),
#               answer('Mode', correct = T),
#               message = 'As a categorical variable, there is no numeric relationship between values in the gender variable. Therefore, 
#               the mode is the only measure of central tendency that is appropriate to describe this variable'),
#      question_numeric("What is the average score on Assessment A?",
#               answer(39.10, correct=T)),
#      question_numeric("What is the most common age in this sample?",
#               answer(19, correct=T)),
#      question_numeric("What is the median score on Assessment D?",
#               answer(81.87, correct=T))
# )
```

<details><summary><span style = "font-weight: bold; font-size: 16pt"> + Click here for a hint </span></summary>
Hint here.
</details>
</br>

<details><summary><span style = "font-weight: bold; font-size: 16pt"> + Click here for the solution </span></summary>

Solution here.

</details>
</br>

## Run Post-Hoc Analyses

Task here.

```{r q7, echo = FALSE}
# quiz(caption = 'Test Your Understanding',
#      question_numeric("What is the IQR for Assessment B?",
#               answer(4, correct=T),
#               message = "To get the IQR for a variable, you'll need to subtract the 25th percentile from the 75th percentile."),
#      question_numeric("What is the maximum score on Assessment C?",
#               answer(78, correct=T)),
#      question_numeric("What is the standard deviation of scores on Assessment D? Please report your answer to two decimal places.",
#               answer(2.15, correct=T))
# )
```

<details><summary><span style = "font-weight: bold; font-size: 16pt"> + Click here for a hint </span></summary>

Hint here.

</details>
</br>

<details><summary><span style = "font-weight: bold; font-size: 16pt"> + Click here for the solution </span></summary>

Solution here.

</details>
</br>

## Calculate Effect Sizes

Task here.

```{r distPlot, echo = F}
# 
# fluidRow(column(4, 
#   sliderInput("mean","Value of Mean:", min = -6, max = 6, value = 0, step = 2),
#   sliderInput('sd', 'Value of SD:', min = 1, max = 5, value = 1, step = .25)),
#   column(8, plotOutput('distPlot')))
# 
# tags$style(HTML(".js-irs-0 .irs-single, .js-irs-0 .irs-bar-edge, .js-irs-0 .irs-bar {background: #4CA384}"))
# tags$style(HTML(".js-irs-1 .irs-single, .js-irs-1 .irs-bar-edge, .js-irs-1 .irs-bar {background: #4CA384}"))

```

<details><summary><span style = "font-weight: bold; font-size: 16pt"> + Click here for a hint </span></summary>

Hint here.

</details>
</br>

<details><summary><span style = "font-weight: bold; font-size: 16pt"> + Click here for the solution </span></summary>

Solution here.

</details>
</br>


## Visualise the Analysis Results

Task here.

<details><summary><span style = "font-weight: bold; font-size: 16pt"> + Click here for a hint </span></summary>

Hint here.

</details>
</br>

<details><summary><span style = "font-weight: bold; font-size: 16pt"> + Click here for the solution </span></summary>

Solution here.

</details>
</br>

## Report & Interpret

Task here.


<details><summary><span style = "font-weight: bold; font-size: 16pt"> + Click here for a hint </span></summary>

Hint here.

</details>
</br>

<details><summary><span style = "font-weight: bold; font-size: 16pt"> + Click here for the solution </span></summary>

Solution here.

</details>
</br>

