---
title: "Week 3 Lab: Intro to NHST"
output:
  learnr::tutorial:
    progressive: true
    allow_skip: true
    css: css/style.css
runtime: shiny_prerendered
---

Make a shiny plot that lets them explore how the AUC works with a probability distribution.


```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(kableExtra)
library(knitr)
library(shiny)
library(shinyjs)
library(bslib)

baseColor <- '#4CA384'
accent1 <- '#9AD079'
accent2 <- '#C4C6C7'
accent3 <- '#19424C'
#tutorial_options(exercise.timelimit = 60)
```

```{r, context = 'server'}
output$distPlot <- renderPlot({
  set.seed(806)
      dat <- data.frame(X=rnorm(300, mean = input$mean, sd = input$sd))
      ggplot(data.frame(X = c(-30, 30)), aes(x = X)) +
        geom_histogram(data=dat, aes(x=X, y = after_stat(density)), fill = '#4CA384', color = '#19424C',
                       binwidth = 2) +
        stat_function(fun=dnorm, geom = "line", args = list(mean = input$mean, sd = input$sd),
                      linewidth = 1, color = '#19424C') +
        scale_x_continuous(breaks = seq(-30, 30, 10), limits=c(-30, 30)) +
        scale_y_continuous(limits=c(0, .4)) +
        theme(axis.title.x = element_text(size = 16, face = 'bold'),
              axis.text.x = element_text(size = 16),
              axis.line.y = element_blank(),
              axis.text.y = element_blank(),
              axis.ticks = element_blank(),
              axis.title.y = element_blank())
    })
```

## Intro to Today's Lab


Intro Goes Here   

```{r, echo = F}
# dat <- read.csv('Week1LabData.csv')
# datInfo <- data.frame(VariableName=colnames(dat), 
#                       Description=c('Participant ID', 'Age in years', 'Time of class in which assessment was given', 'Levels: 1=Male,
#                                     2=Females, 3=Nonbinary', 'Assessment Version A Scores', 'Assessment Version B Scores', 'Assessment 
#                                     Version C Scores', 'Assessment Version D Scores'))
# 
# datInfo %>%
#   kbl(col.names=c('Variable Name', 'Description')) %>%
#   kable_styling(full_width = F) %>%
#   row_spec(0, bold = T, color=baseColor, font_size = 18, align='l') %>%
#   column_spec(1, bold = T, width = '4.5cm')
```


### Learning Objectives
At the end of this lab, you will be able to:

1.  


## Task 1:

Task 1 Here

```{r q1, echo = FALSE}
# quiz(caption = 'Test Your Understanding',
#      question("What kind of file is 'Week1LabData'?",
#               answer("SPSS datafile"),
#               answer("comma separated values file", correct = T),
#               answer("Excel file"),
#               answer("Word file"),
#               message = "The .csv file extension indicates the file is a comma separated values file. This filetype is meant for
#               spreadsheet-style data. It can be easily used with a wide variety of data analysis software, making it more flexible than
#               SPSS (.sav) or Excel (.xlsx) files. Word files (.docx) are not meant to be used with spreadsheet-style data."),
#      question('If you were to restart SPSS, you would open "Week1LabData_SPSS" using "File>Import".',
#               answer('True'),
#               answer('False', correct = T),
#               message = 'Because "Week1LabData_SPSS" is a .sav file, it can be opened using "File>Open". Using the Import option is only
#               necessary when opening a file that is not an SPSS datafile, such as a .csv or .xlsx file.')
#      )
```

<details><summary><span style = "font-weight: bold; font-size: 16pt"> + Click here for a hint </span></summary>

Hint Goes Here

</details>

</br>

<details><summary><span style = "font-weight: bold; font-size: 16pt"> + Click here for the solution </span></summary>

Solution Goes Here

</details>

</br>

## Task 2: 

Task Goes Here

```{r q2, echo = FALSE}
# quiz(caption = 'Test Your Understanding',
#      question("Which type of data is C?",
#               answer("Categorical"),
#               answer("Continuous", correct = T),
#               random_answer_order = T),
#      question("Which type of data is Gender?",
#               answer("Continuous"),
#               answer('Categorical', correct = T)),
#      question("Which type of data is ClassTime?",
#               answer("Continuous"),
#               answer('Categorical', correct = T),
#               message='Categorical data are nominal or ordinal data with no numeric relationship (although they may be represented
#               numerically), while continuous data have a traditional numeric relationship; continuous data can technically take on an  
#               infinite range of values if measurement was precise enough'),
#      question('Which of the following "Measure" values used by SPSS reflects categorical data?',
#               answer('Nominal', correct = T),
#               answer('Ordinal', correct = T),
#               answer('Scale'),
#               message = 'Scale can be used for both types of continuous data (ratio and interval scales of measurement), but categorical
#               variables are split into nominal and ordinal measures. Remember that nominal data reflects purely independent categories,
#               but ordinal data has some sort of implied order (e.g. Likert Scale data, level of education).'))
```

<details><summary><span style = "font-weight: bold; font-size: 16pt"> + Click here for a hint </span></summary>

Hint Here

</details>

</br>

<details><summary><span style = "font-weight: bold; font-size: 16pt"> + Click here for the solution </span></summary>

Solution here.

</details>

</br>


## Task 3: 

Task Description Here

<details><summary><span style = "font-weight: bold; font-size: 16pt"> + Click here for a hint </span></summary>

Hint Here

</details>

</br>

<details><summary><span style = "font-weight: bold; font-size: 16pt"> + Click here for the solution </span></summary>

Solution Here.

</details>

</br>




## Task 4: 

Task Here.

```{r q4, echo = FALSE}
# quiz(caption = 'Test Your Understanding',
#      question("At least one of the groups is limited in size compared to the others",
#               answer("True", correct = T),
#               answer("False"),
#               message = 'The nonbinary group is much smaller than the other two measured groups.'),
#      question_numeric("How many total datapoints do you have for the gender variable?",
#               answer(48, correct=T)),
#      question_numeric("What percentage of participants are female?",
#               answer(41.7, correct=T)),
#      question_numeric('How many nonbinary participants are in your sample?',
#               answer(2, correct=T)))
```

<details><summary><span style = "font-weight: bold; font-size: 16pt"> + Click here for a hint </span></summary>

Hint Here

</details>
</br>

<details><summary><span style = "font-weight: bold; font-size: 16pt"> + Click here for the solution </span></summary>

Solution Here

</details>
</br>

## Task 5: 

Task Here

<details><summary><span style = "font-weight: bold; font-size: 16pt"> + Click here for a hint </span></summary>

Hint Here.

</details>
</br>

<details><summary><span style = "font-weight: bold; font-size: 16pt"> + Click here for the solution </span></summary>

Solution Here.

</details>
</br>

## Task 6: 

Task Description.

```{r q6, echo = FALSE}
# quiz(caption = 'Test Your Understanding',
#      question("Which of the following central tendency measures could be used to describe age? Please select all that apply.",
#               answer("Mean", correct = T),
#               answer("Median", correct = T),
#               answer('Mode', correct = T),
#               message = 'All of the above could technically be used to describe the Age variable. However, it is most common to use the
#               mean, as it includes information from each observation.'),
#      question("Which of the following central tendency measures could be used to describe gender? Please select all that apply.",
#               answer("Mean"),
#               answer("Median"),
#               answer('Mode', correct = T),
#               message = 'As a categorical variable, there is no numeric relationship between values in the gender variable. Therefore, 
#               the mode is the only measure of central tendency that is appropriate to describe this variable'),
#      question_numeric("What is the average score on Assessment A?",
#               answer(39.10, correct=T)),
#      question_numeric("What is the most common age in this sample?",
#               answer(19, correct=T)),
#      question_numeric("What is the median score on Assessment D?",
#               answer(81.87, correct=T))
# )
```

<details><summary><span style = "font-weight: bold; font-size: 16pt"> + Click here for a hint </span></summary>
Hint here.
</details>
</br>

<details><summary><span style = "font-weight: bold; font-size: 16pt"> + Click here for the solution </span></summary>

Solution here.

</details>
</br>

## Task 7: 

Task here.

```{r q7, echo = FALSE}
# quiz(caption = 'Test Your Understanding',
#      question_numeric("What is the IQR for Assessment B?",
#               answer(4, correct=T),
#               message = "To get the IQR for a variable, you'll need to subtract the 25th percentile from the 75th percentile."),
#      question_numeric("What is the maximum score on Assessment C?",
#               answer(78, correct=T)),
#      question_numeric("What is the standard deviation of scores on Assessment D? Please report your answer to two decimal places.",
#               answer(2.15, correct=T))
# )
```

<details><summary><span style = "font-weight: bold; font-size: 16pt"> + Click here for a hint </span></summary>

Hint here.

</details>
</br>

<details><summary><span style = "font-weight: bold; font-size: 16pt"> + Click here for the solution </span></summary>

Solution here.

</details>
</br>

## Task 8: 

Task here.

```{r distPlot, echo = F}
# 
# fluidRow(column(4, 
#   sliderInput("mean","Value of Mean:", min = -6, max = 6, value = 0, step = 2),
#   sliderInput('sd', 'Value of SD:', min = 1, max = 5, value = 1, step = .25)),
#   column(8, plotOutput('distPlot')))
# 
# tags$style(HTML(".js-irs-0 .irs-single, .js-irs-0 .irs-bar-edge, .js-irs-0 .irs-bar {background: #4CA384}"))
# tags$style(HTML(".js-irs-1 .irs-single, .js-irs-1 .irs-bar-edge, .js-irs-1 .irs-bar {background: #4CA384}"))

```

<details><summary><span style = "font-weight: bold; font-size: 16pt"> + Click here for a hint </span></summary>

Hint here.

</details>
</br>

<details><summary><span style = "font-weight: bold; font-size: 16pt"> + Click here for the solution </span></summary>

Solution here.

</details>
</br>


## Task 9: 

Task here.

<details><summary><span style = "font-weight: bold; font-size: 16pt"> + Click here for a hint </span></summary>

Hint here.

</details>
</br>

<details><summary><span style = "font-weight: bold; font-size: 16pt"> + Click here for the solution </span></summary>

Solution here.

</details>
</br>

## Task 10: 

Task here.


<details><summary><span style = "font-weight: bold; font-size: 16pt"> + Click here for a hint </span></summary>

Hint here.

</details>
</br>

<details><summary><span style = "font-weight: bold; font-size: 16pt"> + Click here for the solution </span></summary>

Solution here.

</details>
</br>

