---
title: "Week 2: Intro to NHST"
output:
  learnr::tutorial:
    progressive: true
    allow_skip: true
    css: https://mtruelovehill.github.io/PRM/Labs/css/style.css
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)

pacman::p_load('tidyverse', 'rstatix', 'shiny')

baseColor <- '#4CA384'
accent1 <- '#9AD079'
accent2 <- '#C4C6C7'
accent3 <- '#19424C'

dat <- read.csv('https://mtruelovehill.github.io/PRM/Data/StudentData.csv')
studentDat <- dat[dat$Role=='Student',]
studentDat$Cperc <- (studentDat$C/50)*100
Cdens <- data.frame(cPerc=density(studentDat$Cperc, from = 0, to = 100)$x, dens=density(studentDat$Cperc, from = 0, to = 100)$y)


```

```{r, context = 'server'}
output$aucPlot <- renderPlot({
  loLim <- ifelse(input$auc[1]==0, .00001, input$auc[1])
  hiLim <- ifelse(input$auc[2]==1, .999999, input$auc[2])
  
  ggplot(data.frame(x=c(-6, 6)), aes(x)) + 
    stat_function(fun=dnorm, geom = "line", linewidth = 1.5) +
    geom_area(stat = "function", fun = dnorm, fill = accent4, xlim = c(qnorm(loLim), qnorm(hiLim)),
              alpha = .7) + 
    theme(axis.text = element_blank(),
          axis.ticks = element_blank(),
          axis.title = element_blank()) +
    annotate('text', label = paste0((input$auc[2]-input$auc[1])*100, '% of observations'),
             x = -3, y = .3, size = 6)
    })



output$aucScorePlot <- renderPlot({
  loLim <- ifelse(input$score[1]==31, 31, input$score[1])
  hiLim <- ifelse(input$score[2]==84, 84, input$score[2])


  propVal <- round((pnorm(input$score[2], mean = 58, sd = 9.4)-pnorm(input$score[1], mean = 58, sd = 9.4))*100)

  ggplot(data.frame(x=c(31, 84)), aes(x)) +
    stat_function(fun=dnorm, geom = "line", linewidth = 1.5, args = c(mean = 58, sd = 9)) +
    geom_area(stat = "function", fun = dnorm, fill = accent4, args = c(mean = 58, sd = 9), xlim = c(loLim, hiLim),
              alpha = .7) +
    theme(axis.text = element_blank(),
          axis.ticks = element_blank(),
          axis.title = element_blank()) +
    annotate('text', label = paste0(propVal, '% of observations'), x = 42, y = .03, size = 6)
    })


output$alphaPlot <- renderPlot({
  ggplot(data.frame(x=c(-5, 5)), aes(x)) +
    stat_function(fun=dnorm, geom = "line", linewidth = 1.5, args = c(mean = 0, sd = 1)) + 
    geom_area(stat = "function", fun = dnorm, fill = accent4, args = c(mean = 0, sd = 1),
              xlim = c(qnorm(1-(as.numeric(input$alpha)/2)), 5), alpha = .7) +
    geom_area(stat = "function", fun = dnorm, fill = accent4, args = c(mean = 0, sd = 1),
              xlim = c(-5, qnorm(1-(as.numeric(input$alpha)/2), lower.tail = F)), alpha = .7) +
    scale_x_continuous(breaks = seq(-5, 5, by = .5)) +
    xlab('test statistic') +
    geom_vline(xintercept = qnorm(1-(as.numeric(input$alpha)/2), lower.tail = F), color = accent4, linewidth = 1.5, linetype='dashed') +
    geom_vline(xintercept = qnorm(1-(as.numeric(input$alpha)/2)), color = accent3, linewidth = 1.5, linetype='dashed') +
    theme(axis.text.y = element_blank(),
          axis.ticks.y = element_blank(),
          axis.title.y = element_blank(),
          axis.text.x = element_text(size = 14, face = 'bold'),
          axis.title.x = element_text(size = 16, face = 'bold'))
})

output$pPlot <- renderPlot({
  lowProb <- round(pnorm(input$testStatistic, mean = 0, sd = 1), 4)*100
  hiProb <- round(pnorm(input$testStatistic, mean = 0, sd = 1, lower.tail = F),4)*100

  limVal <- ifelse(input$testStatistic<0, input$testStatistic, input$testStatistic*-1)

  pVal <- ifelse(input$testStatistic<0, round(pnorm(input$testStatistic, mean = 0, sd = 1), 3)*2,
                 round(pnorm(input$testStatistic, mean = 0, sd = 1, lower.tail = F), 3)*2)

  ggplot(data.frame(x=c(-5, 5)), aes(x)) +
    geom_area(stat = "function", fun = dnorm, fill = accent4, args = c(mean = 0, sd = 1),
              xlim = c(-5, limVal), alpha = .7) +
    geom_area(stat = "function", fun = dnorm, fill = accent4, args = c(mean = 0, sd = 1),
              xlim = c(abs(limVal), 5), alpha = .7) +
    stat_function(fun=dnorm, geom = "line", linewidth = 1.5, args = c(mean = 0, sd = 1)) +
    scale_x_continuous(breaks = seq(-5, 5, by = 1)) +
    scale_y_continuous(limits=c(0, .4)) +
    xlab('test statistic') +
    geom_vline(xintercept = input$testStatistic, color = accent3, linewidth = 1.5, linetype='dashed') +
    theme(axis.text.y = element_blank(),
          axis.ticks.y = element_blank(),
          axis.title.y = element_blank(),
          axis.text.x = element_text(size = 14, face = 'bold'),
          axis.title.x = element_text(size = 16, face = 'bold')) +
    geom_rect(aes(xmin = -3.8, xmax = -2.7, ymin = .03, ymax = .07), fill = accent4, alpha = .5) +
    geom_rect(aes(xmin = 3.8, xmax = 2.7, ymin = .03, ymax = .07), fill = accent4, alpha = .5) +
    annotate('text', label = paste0('P', ifelse(pVal==0, ' < .001', paste0(' = ', pVal/2))), x = 3.25, y = .05, size = 6) +
    annotate('text', label = paste0('P', ifelse(pVal==0, ' < .001', paste0(' = ', pVal/2))), x = -3.25, y = .05, size = 6)
})


output$testPlot <- renderPlot({
  limVal <- ifelse(input$testStat2<0, input$testStat2, input$testStat2*-1)
  
  pVal <- ifelse(input$testStat2<0, round(pnorm(input$testStat2, mean = 0, sd = 1), 3)*2, 
                 round(pnorm(input$testStat2, mean = 0, sd = 1, lower.tail = F), 3)*2)
  
  ggplot(data.frame(x=c(-5, 5)), aes(x)) +
    geom_area(stat = "function", fun = dnorm, fill = accent4, args = c(mean = 0, sd = 1),
              xlim = c(-5, limVal), alpha = .7) +
    geom_area(stat = "function", fun = dnorm, fill = accent4, args = c(mean = 0, sd = 1),
              xlim = c(abs(limVal), 5), alpha = .7) +
    stat_function(fun=dnorm, geom = "line", linewidth = 1.5, args = c(mean = 0, sd = 1)) +
    scale_x_continuous(breaks = seq(-5, 5, by = 1)) +
    xlab('test statistic') +
    geom_vline(xintercept = input$testStat2, color = accent3, linewidth = 1.5, linetype='dashed') +
    theme(axis.text.y = element_blank(),
          axis.ticks.y = element_blank(),
          axis.title.y = element_blank(),
          axis.text.x = element_text(size = 14, face = 'bold'),
          axis.title.x = element_text(size = 16, face = 'bold')) +
    annotate('text', label = paste0('p', ifelse(pVal==0, ' < .001', paste0(' = ', pVal))), x = 3.25, y = .2, size = 6)
})
  
```


```{r, echo = F}
tags$style(HTML(".js-irs-0 .irs-from, .js-irs-0 .irs-to, .js-irs-0 .irs-bar {background: #008290; border-color: transparent}"))
tags$style(HTML(".js-irs-1 .irs-from, .js-irs-1 .irs-to, .js-irs-1 .irs-bar {background: #008290; border-color: transparent}"))
tags$style(HTML(".js-irs-2 .irs-single, .js-irs-2 .irs-bar {background: #008290; border-color: transparent}"))
tags$style(HTML(".js-irs-3 .irs-single, .js-irs-3 .irs-bar {background: #008290; border-color: transparent}"))
```

## Intro to Today's Lab

During today's lab, you'll apply the concepts discussed during this week's lecture. Each lab consists of a range of tasks, with corresponding questions you can answer. Please note that the questions are not required and not marked, although they do provide a helpful source of formative feedback that will help you gauge your understanding. 

Today, you'll be using interactive visualisations to demonstrate the concepts covered in the lecture. You will not be using SPSS. 

### Learning Objectives
At the end of this lab, you will be able to:

1. Understand the link between probability and inferential statistics
2. Define statistical terms such as $p$-values and $\alpha$ 
3. Distinguish between one-tailed and two-tailed tests

## Computing the Area Under the Curve

Recall from the [lecture](https://mtruelovehill.github.io/ISAP/Lectures/Week01_IntroToNHST_lecture) that the total area under the curve (AUC) allows us to compute the probability of an observation falling within a given range. Use the plot below to visualise how the AUC corresponds to different proportions of observations in a normal distribution.

```{r aucPlot, echo = F}

sliderInput('auc', min=0, max = 1, step = .01, value = c(0,1), ticks = F, label = NULL)
plotOutput('aucPlot')

```
<br>

Once you have a grasp on what the area under the curve corresponds to, apply this to an example.

Imagine that you have a sample of students who completed an assessment. The plot below reflects the probability distribution of scores across all students who took the assessment. 

```{r, echo = F}
sliderInput('score', 'Assessment Score:', min=31, max = 84, step = 1, value = c(31,84), ticks = F)
plotOutput('aucScorePlot')
```


```{r q1, echo = FALSE}

quiz(caption = 'Test Your Understanding',
     question_numeric('What was the maximum score among students who took the assessment?',
                      answer(84, correct = T),
                      message = 'The maximum score was 84'),
     question_numeric('What percentage of students recieved an A (70+) on the assessment?',
                      answer(10, correct = T),
                      message = '10% of the sample scored 70 or above on the assessment.'),
     question_numeric('If we were to randomly select an exam booklet from the entire pile of assessments, what is the probability that the mark will be between 55 and 65?',
                      answer(40, correct = T),
                      answer(.40, correct = T),
                      message = 'The probability that we the mark will be between 55 and 65 is .40, or 40%'),
     question_numeric('The instructor wants to identify the students who scored in the lowest 5% so that they can check in and offer additional help. Fill in the blank: Students scoring __ and below will be contacted by the instructor.',
                      answer(43, correct = T),
                      message = '5% of the sample scored 43 or below.'),
     question_numeric('Assume this distribution is perfectly normal. What was the mean score on the assessment?',
                      answer(58, correct = T),
                      message = 'In a normal distribution, all observations are centred around the mean. The mean score on this assessment is 58, as 50% of students scored 58 or above, and 50% of students scored 58 or below.')
)
```


## The Null Distribution

When we run an inferential statistical test, it produces a test statistic. The null distribution for that test is the distribution of test statistics we would expect if the null hypothesis were true. If we run a statistical test and our statistic falls in the extreme region of the null distribution, we consider our results to be significant and reject the null hypothesis (because our results would be really unusual if the null hypothesis were true!). See the [lecture](https://mtruelovehill.github.io/ISAP/Lectures/Week01_IntroToNHST_lecture) if you need a refresher. 

Before running a test, we use $\alpha$ to determine our cutoff for significance. $\alpha$ refers to the proportion of values in the null distribution we consider to be extreme. In the plot below, the shaded region indicates the 'extreme' region associated with each value of $\alpha$. 


```{r, echo = F}
fluidRow(column(width = 3, 
                selectInput('alpha', 'Choose an alpha value:', choices = list('.05' = .05, '.01' = .01, '.001' = .001))))

plotOutput('alphaPlot')
```


```{r q2, echo = FALSE}

quiz(caption = 'Test Your Understanding',
     question('Given an alpha of .05, would a test statistic of -1.5 be significant?',
              answer('Yes'),
              answer('No', correct = T),
              message = 'A value of -1.5 would not fall past the alpha threshold, so would not be considered significant.'),
     question('Given an alpha of .001, would a test statistic of 3.5 be significant?',
              answer('Yes', correct = T),
              answer('No'),
              message = 'A value of 3.5 would fall past the alpha threshold and would be considered significant.'),
     question('Given an alpha of .01, would a test statistic of 2.5 be significant?',
              answer('Yes'),
              answer('No', correct = T),
              message = 'A value of 2.5 would not fall past the alpha threshold, so would not be considered significant.')
)
```


## Computing <i> p </i>

Recall from the [lecture](https://mtruelovehill.github.io/ISAP/Lectures/Week01_IntroToNHST_lecture.html) that a $p$-value is the probability of obtaining a statistical test value at least as extreme as the value observed, given that the null hypothesis is true. 

When conducting a one-tailed statistical test, the $p$-value indicates the probability of a value **greater** or **smaller** than the observed value. This means that the $p$-value of a one-tailed test is the probability value from the corresponding tail of the distribution. When conducting a two-tailed statistical test, the $p$-value indicates the probability of a value **as or more extreme than** the observed value. The $p$-value of a two tailed test is calculated by summing the probability of scores more extreme than the observed value on *both* sides of the distribution. 

The plot below shows the probability of each shaded region, as well as a line representing the test statistic value. Use this plot to answer the following questions.

```{r, echo = F}
sliderInput('testStatistic', 'Value of Test Statistic:', min=-5, max = 5, step = .01, value = 0, ticks = F)
plotOutput('pPlot')
```


```{r q3, echo = FALSE}
quiz(caption = 'Test Your Understanding',
     question_numeric("You are conducting a one-tailed test. What is the p-value associated with obtaining a test statistic larger than 1.36? Please report your answer to 3 decimal places.",
              answer(.087, correct = T),
              message = "For a test statistic of 1.36, p = .087. Because it is a one-tailed test (you're looking at the probability of a value larger than the one observed), you only need to check the probability associated with the upper tail of the distribution."),
     question_numeric("You are conducting a one-tailed test. What is the p-value associated with obtaining a test statistic less than -0.93? Please report your answer to 3 decimal places.",
               answer(.176, correct = T),
              message = "For a test statistic of -0.93, p = .176. Because it is a one-tailed test (you're looking at the probability of a value smaller than the one observed), you only need to check the probability associated with the lower tail of the distribution."),
      question_numeric("You are conducting a two-tailed test. What is the p-value associated with obtaining a test statistic as extreme as 2.87? Please report your answer to 3 decimal places.",
              answer(.004, correct = T),
              message = "For a test statistic of 2.87, p = .004 Because you are conducting a two-tailed test, you need the probability of values as extreme as yours (regardless of whether they are positive or negative). You must add the probabilities in each tail. .002 + .002 = .004"),
     question_numeric('Which test statistic reflects the cutoff for the most extreme 5% of the data?',
              answer(-1.96, correct = T),
              answer(1.96, correct = T),
              message = "Because 5% is split between each tail of the distribution, any statistic with a corresponding probability of 2.5% or less would fall into the most extreme 5% of the values. The cutoff point is 1.96."),
     question('Which of the following test statistics are in the most extreme 1% of the data? Please select all that apply.',
              answer(2.76, correct = T),
              answer(2.34),
              answer(-1.74),
              answer(-2.58, correct = T),
              message = "Because 1% is split between each tail of the distribution, any statistic with a corresponding probability of .5% or less would fall into the most extreme 1% of the values.")
)
```


## Putting it all Together

In the last few tasks, we've broken down the general steps of an inferential statistical test. This gives you a glimpse of what's going on under the hood of SPSS. Now, let's put the steps together. 


Use the plot below to answer the following questions. The shaded area reflects $p$, the probability of getting a test statistic at least as extreme as yours given the null is true.

**Important Note:** In the previous section, each capital $P$ reflected the probability of scores within its corresponding region. Here, the lowercase $p$ refers to the $p$-value associated with a ***two-tailed test*** (the total probability of both shaded regions). 

```{r, echo = F}
sliderInput('testStat2', 'Value of Test Statistic:', min=-5, max = 5, step = .01, value = 0, ticks = F)
plotOutput('testPlot')
```



```{r q4, echo = FALSE}
quiz(caption = 'Test Your Understanding',
     question("If alpha = .001 and you're running a two-tailed test, is the p-value associated with a test statistic of 3.06 significant?",
              answer('Yes'),
              answer('No', correct = T),
              message = 'For a test statistic of 3.06, p = .002, which is more than the alpha threshold of .001, so we would not consider this result significant.'),
     question("If alpha = .01 and you're running a two-tailed test, is the p-value associated with a test statistic of -2.75 significant?",
               answer('Yes', correct = T),
              answer('No'),
              message = "For a test statistic of -2.75, p = .006. As this is less than the alpha threshold of .01, this result would be significant."),
     question("If alpha = .05 and you're running a one-tailed test, is the p-value associated with a test statistic of 1.86 significant?",
               answer('Yes', correct = T),
              answer('No'),
              message = "This one was quite tricky. In this plot, the p-value reflects the probability of scores in both tails. To calculate ONLY the probability of scores ABOVE 1.86, you must divide the p-value by two. This would give you the probability of the shaded area in the upper tail. In this case, the p-value would be .062/2, or  .031, which is lower than the alpha threshold of .05. Therefore, this result would be significant.")
)
```




