---
title: "ExtraLectureContent"
author: "Monica Truelove-Hill"
date: "2024-08-27"
output: html_document
---

## Assessing Normality in Data

+ Many of the parametric statistical tests we'll be using are most reliable when data are normally distributed

> **Test Your Understanding:** In a perfectly normal distribution, how are the mean, median, and mode distributed?


---
count: false

## Assessing Normality in Data

+ Many of the parametric statistical tests we'll be using are most reliable when data are normally distributed

+ Not all data are normally distributed
    + **Skewness**

.pull-left[
.center[**Negatively Skewed**]
```{r, echo = F, message = F, fig.height=4}
set.seed(113)
skewDat <- data.frame(x=rbeta(200, 8, 2))
skewDat$x <- skewDat$x*100
skewDat$x <- round(skewDat$x)
skewDat$x[14] <- 85
skewDat$x[27] <- 85

ggplot(skewDat, aes(x)) + geom_histogram(aes(y=after_stat(density)), fill = baseColor, color = accent3, binwidth=2) +
  geom_density(linewidth=1) +
  geom_vline(xintercept = mean(skewDat$x), color = accent1, linewidth = 1.5) +
  geom_vline(xintercept = median(skewDat$x), color = accent2, linewidth = 1.5) +
  geom_vline(xintercept = round(Mode(skewDat$x),2), color = accent3, linewidth = 1.5) +
  annotate('text', label = 'Mean', color = accent1, x = 60, y = .05, size = 6, hjust = 0) +
  annotate('text', label = 'Median', color = accent2, x = 60 , y = .045, size = 6, hjust = 0) +
  annotate('text', label = 'Mode', color = accent3, x = 60, y = .04, size = 6, hjust = 0) +
  theme(axis.text = element_blank(),
        axis.title.y = element_blank(),
        axis.title.x = element_text(size = 14, face = 'bold'),
        axis.ticks = element_blank())

```
]

.pull-right[
.center[**Positively Skewed**]
```{r, echo = F, message = F, fig.height=4}
set.seed(2661)
skewDat <- data.frame(x=rbeta(200, 10, 2))
skewDat$x <- (skewDat$x*100)*-1
skewDat$x <- round(skewDat$x)

ggplot(skewDat, aes(x)) + geom_histogram(aes(y=after_stat(density)), fill = baseColor, color = accent3, binwidth=2) +
  geom_density(linewidth=1) +
  geom_vline(xintercept = mean(skewDat$x), color = accent1, linewidth = 1.5) +
  geom_vline(xintercept = median(skewDat$x), color = accent2, linewidth = 1.5) +
  geom_vline(xintercept = round(Mode(skewDat$x),2), color = accent3, linewidth = 1.5) +
  annotate('text', label = 'Mean', color = accent1, x = -70, y = .05, size = 6, hjust = 0) +
  annotate('text', label = 'Median', color = accent2, x = -70 , y = .045, size = 6, hjust = 0) +
  annotate('text', label = 'Mode', color = accent3, x = -70, y = .04, size = 6, hjust = 0) +
  theme(axis.text = element_blank(),
        axis.title.y = element_blank(),
        axis.title.x = element_text(size = 14, face = 'bold'),
        axis.ticks = element_blank())
```
]

---
## Assessing Normality in Data

+ Many of the parametric statistical tests we'll be using are most reliable when data are normally distributed

+ Not all data are normally distributed
    + **Skewness**
    + **Kurtosis**

.pull-left[
.center[**Platykurtic**]
```{r, echo = F, message = F, fig.height=4}
platDat <- data.frame(x=c(12:20, 20, rep(21, 5), 22, 22, 23, 23, 24:32, 13, 15, 15, 18, 27, 30, 30))

ggplot(platDat, aes(x)) + geom_histogram(aes(y=after_stat(density)), fill = baseColor, color = accent3, binwidth=2) +
  geom_density(linewidth=1, adjust = 1.5) +
  theme(axis.text = element_blank(),
        axis.title.y = element_blank(),
        axis.title.x = element_text(size = 14, face = 'bold'),
        axis.ticks = element_blank())

```
]

.pull-right[
.center[**Leptokurtic**]
```{r, echo = F, message = F, fig.height=4}
leptDat <- data.frame(x = c(12, 13, rep(14, 3), rep(15, 6), rep(16, 25), rep(17, 6), rep(18, 3), 19, 20))

ggplot(leptDat, aes(x)) + geom_histogram(aes(y=after_stat(density)), fill = baseColor, color = accent3, binwidth=1) +
  geom_density(linewidth=1, adjust = 1) +
  theme(axis.text = element_blank(),
        axis.title.y = element_blank(),
        axis.title.x = element_text(size = 14, face = 'bold'),
        axis.ticks = element_blank())
```
]

---
## Assessing Normality in Data

+ Can visually assess histograms

+ Can also assess using statistical tests:

  + For samples < 50, could use the *Shapiro-Wilk Test*

  + For samples > 50, could use the *Kolmogorov–Smirnov Test*

  + Significant results on these tests indicate that normality has been violated

+ Can also check the skewness and kurtosis values
  + Values further from 0 indicate greater levels of skewness or kurtosis